Relocating a cluster of earthquakes using a single station

### 2012-07-20 paper2/paper2.tex
The version sent to Malcolm for comment

%\documentclass[extra,mreferee]{gji}
\documentclass[extra]{gji}

@@

\title{Relocating a cluster of earthquakes using a single station}
\author[D.J. Robinson, M. Sambridge, R. Snieder and J Hauser]
  {David J. Robinson$^{1,2}$, Malcolm Sambridge$^{1}$, Roel Snieder$^{3}$, J\"urg Hauser$^{4}$ \\
  $^1$ Research School of Earth Sciences, Australian National
    University, Canberra ACT \emph{0200}, Australia, E-mail: david.robinson@anu.edu.au \\
$^2$  Risk Analysis Methods, Geoscience Australia, GPO Box \emph{383} Canberra ACT \emph{2601} Australia, E-mail: david.robinson@ga.gov.au\\
$^3$ Center for Wave Phenomena and Department of Geophysics, Colorado School of Mines, Golden CO \emph{80401-1887}, USA \\
$^4$ Earthquakes and the Environment Program, NORSAR,Kjeller, Norway}

@@

\begin{summary}
 Coda wave interferometry (CWI) is an emerging tool for constraining 
earthquake source properties from the intereference pattern of coda waves between nearby
events. A new earthquake location algorithm is derived which relies on CWI based probabilistic estimates 
of earthquake separation. The algorithm can be used with CWI estimates alone or in tandem with travel time
data. Synthetic examples in 2D and 3D and real earthquakes on the 
Calaveras Fault, California are used to demonstrate the usefulness of the CWI technique in locating 
poorly recorded earthquakes. It is demonstrated that CWI: (a) outperforms traditional earthquake location 
techniques when the number of stations is small; (b) is self-consistent across a broad range of station
situations; and (c) can be used with a single station to locate earthquakes.

@@

We design a 2D synthetic experiment (Example 1) by
randomly selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq \hat{x} \leq 50$ and $-50 \leq \hat{y} \leq 50$.
These are indicated with triangles in Fig. \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using eq. (\ref{eq:mu1}). This assumption ensures that the sample
mean of hypothetical separation estimates is consistent with known
CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimisation procedure on the hypothetical CWI data yields the
circles in Fig. \ref{fig-2D50eq-relocation_eg1}. The optimisation
does not lead to the correct solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimisation event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 3 - Synthetic relocation of 50 earthquakes in 2D using all constraints with noise
$\bar{\sigma}_N= 2 \sigma_{1to5Hz}(\delta_t)$.
 Actual and optimisation event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is overly simplistic because we
fix $\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase
the uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where
$\sigma_{1to5Hz}(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig4(b) of ][]{dr_Robinson11a}. Repeating the
optimisation leads to the circles in Fig.
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

We can draw three observations from the error structure in Figs.
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by grey bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors appear to be circular, despite our attempt to
correct for rotational non-uniqueness with the local coordinate
system.

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 2D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The bottom subplot shows the average minimum number of branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

Figure \ref{fig:optimisationresults-2Dsynth}(a) and (b) illustrates 
the maximum $\Delta_{max}$ (top) and mean $\Delta_\mu$ (middle) of the
coordinate difference as a function of increasing linkage. We show
the statistics for the `best' optimisation solution (black) and for
the solution space when all 25 optimisations are considered (grey).
In this case the best solution is determined by the set of event
locations which lead to the smallest value of $L$. The
error in the best solution is consistent when 30\%
or more of the branches are used. The errors increase when only
10\% or 20\% of the constraints are included. Interestingly, this
breakdown around 20\% to 30\% coincides with the point where the
average number of branches required to link each pair reaches 2 (see
Fig.\ref{fig:optimisationresults-2Dsynth} (bottom)). Since the average
number of branches can be computed in advance it can be used as an
indication of the inversion stability prior to
optimisation. An earlier breakdown is observed when all 25 solutions
are considered collectively. For example, the maximum coordinate
error $\Delta_{max}$ exceeds that for the best
solution for linkage $\leq$60\% confirming that the optimisation is susceptible to
local minima and that a range of starting points should
be considered.

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 3D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The absence of the blue and red lines below 60\% and 30\% indicates a breakdown in the solutions
when all or best optimisation result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

When 70\% of the direct constraints are considered all optimisation
results (grey) are consistent with the best solution (black). The
best solution constrains the event locations down to 30\% of the
direct links.

There is one notable difference between the 3D and 2D results. In 2D
the final iteration was typically close to the actual solution when
the optimisation failed to converge.
Contrastingly, in 3D the optimisation appears to converge to the
correct solution or fail completely, leading to a set locations
which in no way resemble the actual solution. This is depicted in
Figure \ref{fig:optimisationresults-3Dsynth} by the absence of the
grey and black lines below 60\% and 30\% of the constraints,
respectively. The reason for this difference may be due to the
increased number of degrees of freedom in 3D requiring a greater
number of iterations to converge. Nevertheless, the accurate
convergence of the best solution for cases with 30\% linkage or
higher is encouraging for the potential of coda wave optimisation to
constrain earthquake location.

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/gmt_california/CaliforniaCalaverasMap1.eps}
\caption{Elevation in California showing location of the 308 event
Calaveras cluster (black star) and 805 seismic stations (red
triangles).} \label{fig:-eqopti-California-Calaveras}
\end{figure}

\begin{table}
\caption{Experiment locations performed for 68 earthquakes on teh
Calaveras Fault} \label{tab:examples}
\begin{tabular}{ll}
\hline
Example 5 & comparison of CWI, catalogue and hypoDD \\
 & locations. \\
Example 6 & exploration of station dependance for CWI and \\
 & hypoDD. \\
Example 7 & combined use of CWI and travel time data \\
& with all and a reduced number of stations. \\
Example 8 & combined use of CWI and travel time data \\
 & when travel times constrain only 50\% of the events. \\
 \hline
\end{tabular}
\end{table}

@@

% Now we have the SVD version
\begin{figure*}
\includegraphics{diags/CalaverasLoc1_hypoDD_SVD.eps}
\caption{Comparison of earthquake hypocenters using three different
methods: catalogue location (column 1), hypoDD with all 308 events
(column 2) and CWI example 1 (column 3). Note that in the case of
the CWI locations we consider only the 68 earthquakes in red, the
blue events are shown for the purpose of orientation only.}
\label{fig-69Calaverasevents_eg1}
\end{figure*}
%%=======================================================================================

@@

CWI locations for the 68 events are illustrated in column two of
Figure \ref{fig-69Calaverasevents_eg1}. Catalogue locations (grey)
are shown for the remaining 240 earthquakes and are included to ease
comparison. The third column of Figure
\ref{fig-69Calaverasevents_eg1} illustrates the locations given by
hypoDD with Singular Value Decomposition (SVD), absolute arrival
times and cross correlation computed time differences.

@@

Accurate location of the Calaveras events is possible using arrival
phases because of the excellent recording situation in California
with many stations and strong azimuthal coverage (see figure
\ref{fig:-eqopti-California-Calaveras}). In contrast, a small number
of stations and poor azimuthal coverage are common limitations when
trying to locate intraplate clusters. For example, there are
only 3 network seismic stations in the South West Seismic Zone of
Western Australia, a region similar in size to that hosting 805
stations in Figure \ref{fig:-eqopti-California-Calaveras}.

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/matlab/Calaveras_substationmap}
\caption{Location of the 10 stations (triangles) used to relocate
the Calaveras events in Example 2. Stations are removed one at a
time according to the order in Table
\ref{tab:Calaveras-stationremoval} and the events relocated. The
events are indicated with black circles.}
\label{fig:-eqopti-Calaveras-substations}
\end{figure}

@@

\begin{figure*}
\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc2.eps}
\caption{....} \label{fig-CWIreducesstats}
\end{figure*}

@@

% This is the version using hypoDD with SVD
\begin{figure*}
\includegraphics[height = 25pc]{diags/CalaverasLoc3_hypoDD_SVD.eps}
\caption{HypoDD (SVD) with reduced stations.}
\label{fig-HYPODDreducesstats}
\end{figure*}
%%=========================================================================

@@

% This is the version using hypoDD with SVD
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasLoc4_hypoDD_SVD.eps} \caption{Statistics on
coordinate differences for reduced station inversions. Differences
are computed between the inversion results (CWI and hypoDD) and the
complete hypoDD locations for all 308 events. The top subplot
illustrates the number of constrainable events in the CWI and hypoDD
inversions as a function of the stations considered.}
\label{fig-statremoval_summarystats}
\end{figure}
%%=========================================================================

@@

The remaining rows of Figure \ref{fig-statremoval_summarystats}
illustrate a statistical comparison of the CWI and hypoDD reduced
station locations to those using hypoDD with all available data. For
the CWI inversions the mean and maximum coordinate difference is
consistent regardless of the number of stations considered. In
contrast, the statistics for the hypoDD inversions fluctuate between
... and .... for the mean coordinate error and ... and ...  for the
maximum confirming that the hypoDD inversion procedure is less
stable than CWI with fewer stations.

@@

\subsection{Example 7 - Combining travel time and CWI constraints}
 Figure
\ref{fig-68Calaverasevents_ttandcoda1} illustrates the earthquake
locations obtained when we combine the travel time and coda wave
data using all data (left) and 5 stations (right). The linear
features observed in the original hypoDD inversions (see figure
\ref{fig-69Calaverasevents_eg1}) is evident in both cases. However,
the coda waves introduce a scatter around these streaks. That is,
the locations in figure \ref{fig-68Calaverasevents_ttandcoda1}
result from a trade-off between hypoDD's desire to place the events
on linear features and the coda waves voracity to push them away
from streaks. When all stations are used the hypoDD constraints are
strong and little off-streak scatter is introduced. As we reduce
hypoDD's leverage by decreasing the number of stations to 5, we
observe an increase in off-streak scatter resulting from the
enhanced influence of the coda.

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc5_hypoDD_SVD.eps}
\caption{Combined HypoDD (SVD) and CWI: all stations and 5
stations.} \label{fig-68Calaverasevents_ttandcoda1}
\end{figure}

@@

An estimate of the cummulative number of aftershocks $N(t)$ after 
$t$ days is given by the modified Omori formula
\begin{equation}
N(t) = K \frac{c^{1-p} + (t+c)^{1-p}}{p-1}
\end{equation}
\citep{dr_Utsu95a}.
The empirically derived constants, $K$, $C$ and $p$ vary between
tectonic settings. For example, using recorded aftershocks with
$M\ge3.2$ of the  Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of
12 July 1993, \citet{dr_Utsu95a} obtained maximum likelihood
estimates for $K$, $p$ and $c$ of 906.5, 1.256 and 1.433,
respectively.  With these empirically derived values an array
deployed within 4 days and left for 150 days will record roughly one
half of the aftershocks occurring within the first 1000 days. That
is,
\begin{equation}
\frac{N(150+4)-N(4)}{N(1000)} = \frac{2257-934}{2626} \approx 0.5.
\end{equation}
This idea is illustrated in Figure \ref{fig:Omorifigure} which shows
the best fitting Omori Formula separated into segments before
(dashed), during (solid) and after (dashed) the pseudo temporary
deployment.

\begin{figure}
\noindent\includegraphics[width = 20pc]{diags/OmoriFigure.eps}
\caption{Cumulative number of aftershocks for the
Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of 12 July 1993
according to the best fitting modified Omori Formula
\citep{dr_Utsu95a}. The leftmost dashed, middle solid and rightmost
dashed signify aftershocks occurring before, during and after the
deployment of a temporary array installed 4 days after the main
shock and left for 150 days. A temporary deployment of this kind
will record roughly 50\% of the aftershocks in the 1000 days
following the mainshock. } \label{fig:Omorifigure}
\end{figure}

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc6_hypoDD_SVD.eps}
\caption{Mimicking the deployment of a temporary network by ignoring
data from all but station CCO for 50\% (or 34) of the events using
by combining CWI and TT (left) and using TT only (right). Only by
combining the data is it possible to locate all 68 events.
Furthermore, combining the data leads to a solution more consistent
with figure ..... } \label{fig-68Calaverasevents_ttsubsetandcoda1}
\end{figure}
%%=========================================================================

@@

Coda wave interferometry is an emerging technique for constraining 
earthquake location. The technique relies on the interference between coda waves
of closely located events and is hence most useful
for studying earthquake clusters and/or aftershock sequences. 
Coda wave constraints are independent of travel times
and can be used in isolation or combination with early onset body waves. The strength of coda
is that it is possible to constrain earthquake location from a single station, 
an outcome demonstrated most clearly by Figures .......

Coda wave interferometry offers a new technique for understanding earthquakes in intraplate areas 
with sparse networks and poor azimuthal coverage.
In particular, the ability to combine coda wave constraints with travel times makes it possible
to link well constrained events from 
a temporary deployment with those recorded outside
the deployment period. All that is required to achieve this is at 
least one network station which has recorded
sufficient events from both periods therefore opening new avenues
for imaging intraplate fault structures and 
improving our understanding of intraplate seismicity and earthquake hazard.

>>>>

%\documentclass[extra,mreferee]{gji}
%\documentclass[extra, onecolumn, doublespacing]{gji}
\documentclass[extra]{gji}

@@

\title{Relocating a cluster of earthquakes using a single station}
\author[D.J. Robinson, M. Sambridge, R. Snieder and J Hauser]
  {David J. Robinson$^{1,2}$, Malcolm Sambridge$^{1}$, Roel Snieder$^{3}$, J\"urg Hauser$^{4}$ \\
  $^1$ Research School of Earth Sciences, Australian National
    University, Canberra ACT \emph{0200}, Australia, E-mail: david.robinson@anu.edu.au \\
$^2$  Risk Analysis Methods, Geoscience Australia, GPO Box \emph{383} Canberra ACT \emph{2601} Australia, E-mail: david.robinson@ga.gov.au\\
$^3$ Center for Wave Phenomena and Department of Geophysics, Colorado School of Mines, Golden CO \emph{80401-1887}, USA \\
$^4$ ..................}

@@

\begin{summary}
 Coda wave interferometry (CWI) is an emerging tool for constraining
earthquake source properties from the intereference pattern of coda
waves between nearby events. A new earthquake location algorithm is
derived which relies on CWI based probabilistic estimates of
earthquake separation. The algorithm can be used with CWI estimates
alone or in tandem with travel time data. Synthetic examples in 2D
and 3D and real earthquakes on the Calaveras Fault, California are
used to demonstrate the potential of CWI for locating poorly
recorded earthquakes. It is demonstrated that CWI: (a) outperforms
traditional earthquake location techniques when the number of
stations is small; (b) is self-consistent across a broad range of
station situations; and (c) can be used with a single station to
locate earthquakes.

@@

We design a 2D synthetic experiment (Example 1) by randomly
selecting $\hat{x}$- and $\hat{y}$-coordinates such that $-50 \leq
\hat{x} \leq 50$ and $-50 \leq \hat{y} \leq 50$. These are indicated
with triangles in Figure \ref{fig-2D50eq-relocation_eg1}.
 We assume a local velocity of $v=3300\,$ms$^{-1}$ between all
event pairs and a dominant frequency of 2.5$\,$Hz to represent waveform data filtered between 1 and 5$\,$Hz.
The CWI data are defined by the positive bounded Gaussian with statistics
$\bar{\mu}_N$ and $\bar{\sigma}_N$. A hypothetical CWI mean is created by setting
\begin{equation}
\label{eq:hypothetical-CWI-mean-optichapt}
\bar{\mu}_N = \mu_1 \left( \widetilde{\delta}_t \right)
\end{equation}
using eq. (\ref{eq:mu1}). This assumption ensures that the sample
mean of hypothetical separation estimates is consistent with known
CWI biases \citep{dr_Robinson11a}. In example 1 we use
$\bar{\sigma}_N=0.02$ between all event pairs. Application of our
optimisation procedure on the hypothetical CWI data yields the
circles in Figure \ref{fig-2D50eq-relocation_eg1}. The optimisation
does not lead to the correct solution due to the addition of noise
($\bar{\sigma}_N=0.02$) on the hypothetical CWI data. The average
coordinate error is 2.0\,m which is small compared to the noise of
$\bar{\sigma}_N=0.02$ which for $v_s=3300$\,ms$^{-1}$ and
$f_{dom}=2.5$\,Hz corresponds to roughly 25\,m.
\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_1.eps}
\caption{Example 1 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N=0.02$. Actual and
optimisation event locations are shown in triangles and circles,
respectively.} \label{fig-2D50eq-relocation_eg1}
\end{figure}

\begin{figure}
\includegraphics[width = 20pc]{diags/locs_2D_50eq_3.eps}
\caption{Example 2 - Synthetic relocation of 50 earthquakes in 2D
using all constraints with noise $\bar{\sigma}_N= 2
\sigma_{1to5Hz}(\delta_t)$.
 Actual and optimisation event locations
are shown in triangles and circles, respectively.}
\label{fig-2D50eq-relocation_eg3}
\end{figure}
\citet{dr_Robinson11a} demonstrates that the noise on CWI estimates
is often larger than 0.02 and that it increases with event
separation. Consequently, example 1 is overly simplistic because we
fix $\bar{\sigma}_N=0.02$ for all pairs. In example 2 we increase
the uncertainty and introduce a distance dependance into the
hypothetical $\bar{\sigma}_N$ by defining
$\bar{\sigma}_N=\sigma_{1to5Hz}(\delta_t)$, where
$\sigma_{1to5Hz}(\delta_t)$ is the half-width of the errorbars for a
synthetic acoustic experiment with filtering between 1 and 5\,Hz
\citep[see Fig. 4(b) of ][]{dr_Robinson11a}. Repeating the
optimisation leads to the circles in Figure
\ref{fig-2D50eq-relocation_eg3} which have an average coordinate
error of 2.8\,m.

@@

We can draw three observations from the error structure in Figures
\ref{fig-2D50eq-relocation_eg1} and \ref{fig-2D50eq-relocation_eg3}.
Firstly, the location errors depicted by gray bars increase between
examples 1 and 2 with the introduction of larger noise. Secondly,
the errors are larger for events at greater distances from the
center. This is because events near the center of the cluster are
constrained by links from all angles, whereas those on the outside
are moderated by links from a limited number of directions. This
observation is analogous to problems associated with poor azimuthal
coverage in common triangulation problems such as individual
earthquake location from limited travel time data, or GPS
positioning with few satellites. Our third observation is that the
location errors appear to be circular, despite our attempt to
correct for rotational non-uniqueness with the local coordinate
system.

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width =
20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\caption{Example 3 - Statistical measures of error in the
optimisation solutions for the 2D synthetic cases when all (gray)
and best (black) results are considered. The statistics
$\Delta_{max}$ and $\Delta_\mu$  are the maximum and mean coordinate
error, respectively. The bottom subplot shows the average minimum
number of branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

Figure \ref{fig:optimisationresults-2Dsynth}(a) and (b) illustrates
the maximum $\Delta_{max}$ (top) and mean $\Delta_\mu$ (middle) of
the coordinate difference as a function of increasing linkage. We
show the statistics for the `best' optimisation solution (black) and
for the solution space when all 25 optimisations are considered
(gray). In this case the best solution is determined by the set of
event locations which lead to the smallest value of $L$. The error
in the best solution is consistent when 30\% or more of the branches
are used. The errors increase when only 10\% or 20\% of the
constraints are included. Interestingly, this breakdown around 20\%
to 30\% coincides with the point where the average number of
branches required to link each pair reaches 2 (see
Fig.\ref{fig:optimisationresults-2Dsynth} (bottom)). Since the
average number of branches can be computed in advance it can be used
as an indication of the inversion stability prior to optimisation.
An earlier breakdown is observed when all 25 solutions are
considered collectively. For example, the maximum coordinate error
$\Delta_{max}$ exceeds that for the best solution for linkage
$\leq$60\% confirming that the optimisation is susceptible to local
minima and that a range of starting points should be considered.

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width =
20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\caption{Example 4 - Statistical measures of error in the
optimisation solutions for the 3D synthetic cases when all (gray)
and best (black) results are considered. The statistics
$\Delta_{max}$ and $\Delta_\mu$ are the maximum and mean coordinate
error, respectively. The absence of the gray and black lines below
60\% and 30\% indicates a breakdown in the solutions when all or
best optimisation result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

When 70\% of the direct constraints are considered all optimisation
results (gray) are consistent with the best solution (black). The
best solution constrains the event locations down to 30\% of the
direct links.

There is one notable difference between the 3D and 2D results. In 2D
the final iteration was typically close to the actual solution when
the optimisation failed to converge. Contrastingly, in 3D the
optimisation appears to converge to the correct solution or fail
completely, leading to a set locations which in no way resemble the
actual solution. This is depicted in Figure
\ref{fig:optimisationresults-3Dsynth} by the absence of the gray and
black lines below 60\% and 30\% of the constraints, respectively.
The reason for this difference may be due to the increased number of
degrees of freedom in 3D requiring a greater number of iterations to
converge. Nevertheless, the accurate convergence of the best
solution for cases with 30\% linkage or higher is encouraging for
the potential of coda wave optimisation to constrain earthquake
location.

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/gmt_california/CaliforniaCalaverasMap1.eps}
\caption{Elevation in California showing location of the Calaveras
cluster (white star) and 805 seismic stations (black triangles).}
\label{fig:-eqopti-California-Calaveras}
\end{figure}

\begin{table}
\caption{Location examples for the 68 Calaveras earthquakes.}
\label{tab:examples}
\begin{tabular}{ll}
\hline
Example 5 & comparison of CWI, catalogue and hypoDD \\
 & locations. \\
Example 6 & exploration of station dependance for CWI and \\
 & hypoDD. \\
Example 7 & combined use of CWI and travel time data \\
& with all and a reduced number of stations. \\
Example 8 & combined use of CWI and travel time data \\
 & when travel times constrain only 50\% of the events. \\
 \hline
\end{tabular}
\end{table}

@@

% Now we have the SVD version
\begin{figure*}
\includegraphics{diags/CalaverasLoc1_hypoDD_SVD.eps}
\caption{Example 5 - Comparison of earthquake locations using three
different methods: catalogue location (column 1), CWI (column 2) and
hypoDD (column 3). Note that in the case of the hypoDD and CWI
inversions we consider only the 68 earthquakes in black, the gray
events are shown for the purpose of orientation only.}
\label{fig-69Calaverasevents_eg1}
\end{figure*}
%%=======================================================================================

@@

CWI locations for the 68 events are illustrated in column two of
Figure \ref{fig-69Calaverasevents_eg1}. Catalogue locations (gray)
are shown for the remaining 240 earthquakes and are included to ease
comparison. The third column of Figure
\ref{fig-69Calaverasevents_eg1} illustrates the locations given by
hypoDD with Singular Value Decomposition (SVD), absolute arrival
times and cross correlation computed time differences.

@@

Accurate location of the Calaveras events is possible using arrival
phases because of the excellent recording situation in California
with many stations and strong azimuthal coverage (see Fig.
\ref{fig:-eqopti-California-Calaveras}). In contrast, a small number
of stations and poor azimuthal coverage are common limitations when
trying to locate intraplate clusters. For example, there are only 3
network seismic stations in the South West Seismic Zone of Western
Australia, a region similar in size to that hosting 805 stations in
Figure \ref{fig:-eqopti-California-Calaveras}.

@@

\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasMap/matlab/Calaveras_substationmap}
\caption{Location of the 10 stations (triangles) used to relocate
the Calaveras events in Example 6 to 8. Stations are removed one at
a time according to the order in Table
\ref{tab:Calaveras-stationremoval} and the events relocated. The
events are indicated with black circles.}
\label{fig:-eqopti-Calaveras-substations}
\end{figure}

@@

\begin{figure*}
\includegraphics[angle=90,height = 50pc]{diags/CalaverasLoc2.eps}
\caption{Example 6 - CWI locations with reduced stations.}
\label{fig-CWIreducesstats}
\end{figure*}

@@

% This is the version using hypoDD with SVD
\begin{figure*}
\includegraphics[height = 25pc]{diags/CalaverasLoc3_hypoDD_SVD.eps}
\caption{Example 6 - HypoDD (SVD) locations with reduced stations.}
\label{fig-HYPODDreducesstats}
\end{figure*}
%%=========================================================================

@@

% This is the version using hypoDD with SVD
\begin{figure}
\noindent\includegraphics[width =
20pc]{diags/CalaverasLoc4_hypoDD_SVD.eps} \caption{Example 6 -
Statistics on coordinate differences for reduced station inversions.
Differences are computed between the inversion results (CWI and
hypoDD) and the complete hypoDD locations for all 308 events. The
top subplot illustrates the number of constrainable events in the
CWI and hypoDD inversions as a function of the stations considered.}
\label{fig-statremoval_summarystats}
\end{figure}
%%=========================================================================

@@

The remaining rows of Figure \ref{fig-statremoval_summarystats}
illustrate a statistical comparison of the CWI and hypoDD reduced
station locations to those using hypoDD with all available data. For
the CWI inversions the mean and maximum coordinate difference is
consistent regardless of the number of stations considered. In
contrast, the hypoDD mean and maximum coordinate error fluctuate
above those for CWI confirming that the hypoDD inversion is less
stable than CWI with fewer stations.

@@

\subsection{Example 7 - Combining travel time and CWI constraints}
 Figure
\ref{fig-68Calaverasevents_ttandcoda1} illustrates the earthquake
locations obtained when we combine the travel time and coda wave
data using all data (left) and 5 stations (right). The linear
features observed in the original hypoDD inversions (see Fig.
\ref{fig-69Calaverasevents_eg1}) is evident in both cases. However,
the coda waves introduce a scatter around these streaks. That is,
the locations in figure \ref{fig-68Calaverasevents_ttandcoda1}
result from a trade-off between hypoDD's desire to place the events
on linear features and the coda waves voracity to push them away
from streaks. When all stations are used the hypoDD constraints are
strong and little off-streak scatter is introduced. As we reduce
hypoDD's leverage by decreasing the number of stations to 5, we
observe an increase in off-streak scatter resulting from the
enhanced influence of the coda.

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc5_hypoDD_SVD.eps}
\caption{Example 7 - Combined HypoDD (SVD) and CWI locations using
data form all stations (left) and 5 stations (right).}
\label{fig-68Calaverasevents_ttandcoda1}
\end{figure}

@@

An estimate of the cummulative number of aftershocks $N(t)$ after
$t$ days is given by the modified Omori formula
\begin{equation}
\label{eq:CumOmori}
 N(t) = K \frac{c^{1-p} + (t+c)^{1-p}}{p-1}
\end{equation}
\citep{dr_Utsu95a}.
The empirically derived constants, $K$, $C$ and $p$ vary between
tectonic settings. For example, using recorded aftershocks with
$M\ge3.2$ of the  Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of
12 July 1993, \citet{dr_Utsu95a} obtained maximum likelihood
estimates for $K$, $p$ and $c$ of 906.5, 1.256 and 1.433,
respectively.  With these empirically derived values an array
deployed within 4 days and left for 150 days will record roughly one
half of the aftershocks occurring within the first 1000 days. That
is,
\begin{equation}
\frac{N(150+4)-N(4)}{N(1000)} = \frac{2257-934}{2626} \approx 0.5.
\end{equation}
This idea is illustrated in Figure \ref{fig:Omorifigure} which shows
the best fitting Omori Formula separated into segments before
(gray), during (black) and after (gray) the pseudo temporary
deployment.

\begin{figure}
\noindent\includegraphics[width = 20pc]{diags/OmoriFigure.eps}
\caption{Cumulative number of aftershocks for the
Hokkaido-Nansei-Oki, Japan $M_s=7.8$ earthquake of 12 July 1993
using equation \ref{eq:CumOmori}. The leftmost gray, middle black
and rightmost gray lines signify aftershocks occurring before,
during and after the deployment of a pseudo temporary array
installed 4 days after the main shock and left for 150 days. A
temporary deployment of this kind will record roughly 50\% of the
aftershocks in the 1000 days following the mainshock. }
\label{fig:Omorifigure}
\end{figure}

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc6_hypoDD_SVD.eps}
\caption{Example 8 - Mimicking the deployment of a temporary network
by ignoring data from all but station CCO for 50\% (or 34) of the
events. Locations are shown for the combined CWI and travel time
inversion (left) and the inversion with travel times only (right)
Only by combining the data is it possible to locate all 68 events.
Furthermore, combining the data leads to a solution more consistent
with Figure \ref{fig-69Calaverasevents_eg1}. }
\label{fig-68Calaverasevents_ttsubsetandcoda1}
\end{figure}
%%=========================================================================

@@

Coda wave interferometry is an emerging technique for constraining
earthquake location. The technique relies on the interference
between coda waves of closely located events and is hence useful for
studying earthquake clusters and/or aftershock sequences. Coda wave
constraints are independent of travel times and can be used in
isolation or combination with early onset body waves. The strength
of coda is that it is possible to constrain earthquake location from
a single station, an outcome demonstrated most clearly by Figures
Figures \ref{fig-CWIreducesstats} and
\ref{fig-68Calaverasevents_ttsubsetandcoda1}

Coda wave interferometry offers a new technique for understanding
earthquakes in intraplate areas with sparse networks and poor
azimuthal coverage. In particular, the ability to combine coda wave
constraints with travel times makes it possible to link well
constrained events from a temporary deployment with those recorded
outside the deployment period. All that is required to achieve this
is at least one network station which has recorded sufficient events
from both periods. CWI facilitates the location of poorly recorded
events to an accuracy approaching those recorded during the
temporary deployment and therefore opens new avenues for imaging
intraplate fault structures and improved our understanding of
intraplate seismicity and earthquake hazard.

Another potential application of CWI is in the area of hydraulic
fracturing such as hot rock geothermal projects and/or petroleum
reservoir engineering. Monitoring pumping induced micro earthquakes
is a key step in understanding the migration of fluids in such
reservoirs. There is a trade-off in the ability of surface deployed
networks to locate events which are small and/or deep. Downhole
seismic monitoring is likely to play increasingly important roles in
deep reservoir projects. CWI creates new possibilities to monitor
pumping induced micro earthquakes from fewer boreholes and hence
dramatically reduce the costs of reservoir monitoring at large
depths. It may also be possible to utilize coda for understanding
hazard in tunneled mining operations where the location of deep
tunnels prohibits azimuthal coverage of induced events.
